There are too many people to name that have either directly or indirectly
contributed to this thesis. 

First, I would like to extend my deepest thanks to my supervisor, Dr. Robert
Zee. His direct and indirect leadership has pushed me to become a better
engineer. His advice and guidance has helped me expand my perspective on what
constitutes a good engineering solution to a problem. I am fortunate to be part
the environment that he has created at the Space Flight Laboratory which serves
as an incubator for the next generation of space engineers.  Thank you for
seeing my potential and giving me the opportunity to push my boundaries and
thrive.

I would next like to thank my mission manager Nathan Cole. I always felt as
though Nathan made myself and the other students working under him a priority.
He would always have time to discuss any problem and ensured that my work was
not only productive and useful to the mission but also thoroughly enriching.
With Nathan, I would also like to thank the other members of the POPS team:
Matt, Lukasz, and Erica. This thesis was made possible through your
collaboration, patience, and advice.  

To my friends, you have made this master's program not only possible but two of
the best years of my life. To my cohort: Diana, Andreas, Lukasz, Djordje, Kim,
Connor, Hooman, Arjun, Eric, and Jacob, I have enjoyed every moment learning
and sharing this experience with you. I know you will all excel wherever you
go. To Jordy, since we met on the first day of second year of undergrad, you
have been a constant companion. Thank you for your patience, support, and
friendship. To Mia, thank you for showing me what is possible through talent,
grit, and determination. Your example has pushed me to work harder and be
better.  To my friends from home: Momo, Julie, Ethan, Saymon, Gurjeet, Mazzen,
and Daylen, thank you for helping me keep my sense of community. Your
friendship has served as my bedrock. 

Oftentimes omitted in academics is physical exercise. I would like to thank the
members of the Swole Patrol: Nathan, Rami, Suraj, and Matt. Through mentorship
and example, they have taught me how to make very heavy things feel less heavy.

Lastly and the most importantly of all, I would like to thank Mum and Dad. I
find it difficult to put into words the appreciation I have for your
unconditional support. It feels as though you're always there, always know
what's best and always know what to do. Team Farmer.

